---
title: "Homework 2"
author: "Evan Woelk Balzer"
date: "15/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library(tidyverse)

#Question 1#
#1a
H0:There is no difference in catch tonnage between 2009 and 2010 in the Gulf of Mexico fishery.
HA:The 2010 Gulf of Mexico fishery catch tonnage is significantly lower than the 2009 catch tonnage.

#1b
```{r Question 1 bi}
library(ggplot2)
library(tidyverse)
sc <- read_csv("shrimp_catch.csv")
sc
a<- ggplot(sc, aes(`Two thousand nine`))
brx <- pretty(range(sc$`Two thousand nine`), n = nclass.Sturges(sc$`Two thousand nine`),min.n = 1)
NINE <- a+geom_histogram(breaks = brx, alpha = 0.5, color = "black")+labs(x = "Catches (tonnes)", y = "Count")+theme_classic()
NINE
```


```{r Question 1 bii}
b<- ggplot(sc, aes(`Two thousand ten`))
breax <- pretty(range(sc$`Two thousand ten`), n = nclass.Sturges(sc$`Two thousand ten`),min.n = 1)
TEN <- b+geom_histogram(breaks = brx, alpha = 0.5, color = "black")+labs(x = "Catches (tonnes)", y = "Count")+theme_classic()
TEN
```
c) In this example, the catch data for both years has been measured at the same villages. Accordingly, the data can considered paired, and the appropriate test is the paired t-test. Because we are interested in whether a change has happened, a two-sided test is appropriate.
#1d
```{r Question 1 d}
#getting a vector of catch data#
nin <- sc$`Two thousand nine`
nin

tin <- sc$`Two thousand ten`
tin
PairedTCatches <- t.test(nin, tin, paired = TRUE)
PairedTCatches
```
#1e
```{r Question 1 e}
critshrimp <- qt(0.975, 13)
critshrimp
```
The critical value is 2.160369 and the observed t-value is 6.3808. 

#1f
Since tobs > tcrit, we reject the null hypothesis. We have evidence that catch tonnage in the Gulf of Mexico fishery has changed between 2009 and 2010.

#2a
H0: The mean phosphorous level in Lake Erie is not different than 15μg/L
HA: The mean phosphorous level in Lake Eerie is greater than 15μg/L

#2b
```{r Question 2 b}
EeriePhos <- read_csv("EeriePhos.csv")
EP <- EeriePhos[1:10,1:2]
c<- ggplot(EP, aes(EP$Phosphorous))
epbrx <- pretty(range(EP$Phosphorous), n = nclass.Sturges(EP$Phosphorous),min.n = 1)
ephist <- c+geom_histogram(breaks = epbrx, alpha = 0.5, color = "black")+labs(x = "Phosphorous Level (15μg/L)", y = "Count")+theme_classic()
ephist
```

#2c
We only have data from one set of observations and wish to compare an observed mean to a theoretical one.These conditions meet the criteria for a one-sample t-test. Since we want to know whether the mean is higher than the theoretical mean and have evidence to support that prediction, the test should also be one-sided.

#2d
```{r Question 2 d}
eptest <- t.test(EP$Phosphorous, mu = 15)
eptest
```

#2e
```{r Question 2 e}
qt(0.95, 9)
```
tcrit = 1.833113
tobs = 1.8462

#2f
Since tobs > tcrit, we reject the null hypothesis. There is evidence that the observed mean
phosphorous level (17.3415μg/L) in Lake Erie is significantly higher than the acceptable level (15μg/L).

#Question 3
#3a
H0: There is no difference between the growth rates of farmed salmon with vegetarian and carnivorous diets.
HA: There is a difference between the growth rates of farmed salmon with vegetarian and carnivorous diets

#3b
```{r Question 3 bi}
SalmonDiet <- read_csv("SalmonDiet.csv")
d <- ggplot(SalmonDiet, aes(SalmonDiet$Carnivorous))
SDCbrx <- pretty(range(SalmonDiet$Carnivorous), n = nclass.Sturges(SalmonDiet$Carnivorous),min.n = 1)
SDChist <- d+geom_histogram(breaks = SDCbrx, alpha = 0.5, color = "black")+labs(x = "Growth Rate (% body weight/day)", y = "Count")+theme_classic()
SDChist
```

```{r Question 3 bii}
e <- ggplot(SalmonDiet, aes(SalmonDiet$Vegetarian))
SDVbrx<- pretty(range(SalmonDiet$Vegetarian), n = nclass.Sturges(SalmonDiet$Vegetarian),min.n = 1)
SDVhist <- e+geom_histogram(breaks = SDVbrx, alpha = 0.5, color = "black")+labs(x = "Growth Rate (% body weight/day)", y = "Count")+theme_classic()
SDVhist
```

#3c
Because the samples were not taken from the same population across the trials, these data are not paired. Since we are interested in whether there is a difference between the two groups, the appropriate test is an independent two-sample t-test.

#3d
```{r Question 3 d}
sdtest <- t.test(SalmonDiet$Carnivorous, SalmonDiet$Vegetarian)
sdtest
```

#3e
```{r Question 3 e}
qt(0.95, 27.862)
```
tcrit = 1.701419
tobs = 12.237

#3f
Since tcrit >>> t obs, we reject the null hypothesis. There is evidence that farmed salmon on vegetarian and carnivorous diets have different growth rates.

#Question 4
#4a
H0: Trophic level and PCB poison symptoms are independent
HA: Trophic level and PCB poison symptoms are not independent
  
#4b
```{r Question 4 b}
trophicPCB <- read_csv("PCB.csv")
trophicPCB
f <- ggplot(trophicPCB, aes(Outcome, Total, fill = Group ))
tPCBbar <- f + geom_bar(stat = "identity", position = "dodge")+theme_classic()
tPCBbar
```

#4c
We want to know whether categorical variables (PCB poison symptoms and trophic level) are independent from one another. Accordingly, the appropriate test is a chi-square test of independence.

#4d
```{r Question 4 d}
first_trophic <- c(133, 100, 17)
top_predator <- c(4, 61, 52)
prechiPCB <- data.frame('FT'=first_trophic,'TP'= top_predator)
chisq.test(prechiPCB)
```

#4e
```{r Question 4 e}
qchisq(0.95, 2)
```
X2obs = 115.66
X2crit = 5.991465

#4f
Since X2obs > X2crit, we reject the null hypothesis. We have evidence that PCB poison symptoms and trophic level of organisms in the St. Lawrence River are not independent.